---
title: "Lab: Anomaly detection using the `AnomalyDetection` package"
author: "Hicham Zmarrou"
date: "`r Sys.Date()`"
output:
  html_notebook:
    highlight: pygments
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---
___________________________________________________________________________________________________________


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='img/', echo=TRUE, warning=FALSE, message=FALSE,
                      digits = 3)
```


## Load libraries 

```{r}
library(AnomalyDetection)
library(dygraphs)
library(xts)
library(lubridate)
```

## Look for ALL anomalies
In this case, we want to restrict our potential anomalies to only the last day, hour, week, etc. and focus on the entire time-series.

```{r}
data(raw_data)
res <- AnomalyDetectionTs(raw_data, max_anoms=0.02, direction='both', plot=TRUE)
res$plot
```

Using Dygraphs interactive plot

```{r}
series <- xts(raw_data$count, order.by = raw_data$timestamp, tz = "UTC")

plot <- dygraph(series, main = '0.91% Anomalies (alpha=0.05, direction=both)') %>%
  dyRangeSelector() %>%
  dyAxis("y", 'Count') %>%
  dyOptions(useDataTimezone = TRUE) %>%
  dyRoller(rollPeriod = 1)

for(i in 1:length(res$anoms$timestamp)) plot <- plot %>% dyAnnotation(ymd_hms(res$anoms$timestamp[i]), text = 'O', tooltip = res$anoms$timestamp[i])

plot
```


## Only look for anomalies in the last day
The AnomalyDetection package also has an option where you can only look for anomalies in the last `x` periods (days, weeks, months, etc.). In this case, not only does it only find anomalies only in the specified period, the plot changes as well. The Github link goes this in more depth but basically it cuts off the plot earlier and dims the irrelevant time-series. Of course, this isn't needed for the dygraphs case since you can zoom in/out at your own will. Let s check the examples out:

```{r}
data(raw_data)
res <- AnomalyDetectionTs(raw_data, max_anoms=0.02, direction='both', plot=TRUE, only_last = 'day')
res$plot
```

Using `dygraph`

```{r}
series <- xts(raw_data$count, order.by = raw_data$timestamp, tz = "UTC")
names(series) <- 'Count'

plot <- dygraph(series, main = '1.74% Anomalies (alpha=0.05, direction=both)') %>%
  dyRangeSelector(dateWindow = c("1980-09-29 13:58:00 UTC", "1980-10-05 13:58:00 UTC")) %>%
  dyAxis("y", 'Count') %>%
  dyOptions(useDataTimezone = TRUE) %>%
  dyRoller(rollPeriod = 1) %>%
  dyShading(from = min(index(series)), to = "1980-10-04 13:58:00 UTC")

for(i in 1:length(res$anoms$timestamp)) plot <- plot %>% dyAnnotation(ymd_hms(res$anoms$timestamp[i]), text = 'O', tooltip = res$anoms$timestamp[i])

plot
```

We see its easy to have the dygraph start zoomed in on the same days as the original plot, but we have the ability to easily look further back. 