---
title: "Causal Impact Analysis on VolksWagen Emissions Scandal"
author: "Hicham Zmarrou"
date: "`r Sys.Date()`"
output:
  html_notebook:
    highlight: pygments
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---
___________________________________________________________________________________________________________


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='img/', echo=TRUE, warning=FALSE, message=FALSE,
                      digits = 3)
```


## Introduction 
In this exercise we will analyse we will assess the impact that the VolksWagen Emissions Scandal had on their stock price. To this endwe will use the Adjusted Close price for VolksWagens stock as well as BMW and Allianz Insurance;  the last two will be used as regressors/control time series of the VW series in the second part of the analysis. The Emissions Scandal broke on Friday the 18th of September 2015, so I'm going to collect weekly data from the beginning of 2011 up to current date.



```{r}
options(warn = -1)
#install.packages("tseries")
library(tseries)
#install.packages("ggplot2")
library(ggplot2)
#devtools::install_github("google/CausalImpact")
library(CausalImpact)
```


```{r}
start = "2011-01-03"
  end = "2017-03-20"
quote = "AdjClose"
VolksWagen <- get.hist.quote(instrument = "VOW.DE", start, end, quote, compression = "w")
BMW <- get.hist.quote(instrument = "BMW.DE", start, end, quote, compression = "w")
Allianz <- get.hist.quote(instrument = "ALV.DE", start, end, quote, compression = "w")
series <- cbind(VolksWagen, BMW, Allianz)
```

We then plot the three time series.

```{r}

colnames(series) <- c("VolksWagen", "BMW", "Allianz")
autoplot(series, facet = NULL) + xlab("") + ylab("Adjusted Close Price")

``` 

We need to define the pre- and post-intervention periods (the emission scandal started on the 18th of September 2015)

```{r}
pre.period <- as.Date(c(start, "2015-09-14"))
post.period <- as.Date(c("2015-09-21", end))
```


## A Simple Model
The Causal Impact function needs at least three arguments: data, pre.period and post.period. The easiest way to perform a causal analysis is to provide only the series where the intervention took place as the data input and specify the seasonality frequency in the `model.args` parameter. This is equivalent as specifying a local level model with a seasonality component:

1. please fill the `---`

```
impact_vw <- ---(series[, 1], pre.period, post.period, model.args = list(niter = 1000, nseasons = 52))
plot(---)

```

```
summary(---)
```


## Including controls

We can try to improve our model by supplying one or more control time series so that we're basically performing a regression on our response variable. We will use the `BMW` and `Allianz` stock prices to explain our target series (you may argue that those series - especially BMW - may have been influenced by the scandal as well and that may be true, but certainly at a lower magnitude):


2. fill the `---`


```
impact_vw_reg <- ---(---, pre.period, post.period, model.args = list(niter = 1000, nseasons = 52))
plot(---)

```

```
summary(---)
```

