---
title: "Predicting bike sharing demand with Prophet"
author: "Hicham Zmarrou"
date: "`r Sys.Date()`"
output:
  html_notebook:
    highlight: pygments
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---
___________________________________________________________________________________________________________



```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='img/', echo=TRUE, warning=FALSE, message=FALSE,
                      digits = 3)
```


We give `prophet` try and  use a dataset from a Kaggle competition that took place years ago related to forecast [bike sharing demand](https://www.kaggle.com/c/bike-sharing-demand) in Washington.



## The Data
The data set contains several features, including:

* date-time

* holiday (as a binary variable indicating whether the day is considered a holiday)

* weather conditions (a categorical variable with 4 classes)

* temperature

* humidity

* windspeed

* total number of bike rentals

```{r}
# Libraries
library(tidyverse)
library(lubridate)
library(prophet)
library(forecast)

# Read data
bikes <- read_csv('./data/bikes.csv') %>%
    mutate(datetime = date(datetime))
head(bikes,5)
```




The data set provides hourly rental data spanning two years (2011 and 2012). For the competition, the training set is comprised of the first 19 days of each month, while the test set is the 20th to the end of the month. The objective is to predict the total count of bikes rented during each hour covered by the test set, using only information available prior to the rental period.

Because the data is from a competition, the test set does not contain the total count of bikes rented, so for this experiment it cannot be use to evaluate the model's performance. Another issue is that Prophet works best with daily periodicity data, so I had to transform the data from an hourly to a daily time scale.

From the features above, the ones we are interested in are the date-time, holiday and total number of bike rentals.





