---
title: "Train a Random Forest model"
author: "Hicham Zmarrou"
date: "`r Sys.Date()`"
output:
  html_notebook:
    highlight: pygments
    number_sections: no
    theme: cosmo
    toc: yes
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
---


## Train a Random Forest model

Here you will use the `randomForest()` function from the randomForest package to train a Random Forest classifier to 
predict loan default.

* We already loaded  the dataset named `german.data` from the [Machine Learning Library of the UCI](https://archive.ics.uci.edu/ml/datasets/statlog+(german+credit+data))
* Inspect the description of the data from the [Machine Learning Library of the UCI website](https://archive.ics.uci.edu/ml/datasets/statlog+(german+credit+data)) 
* make a training and test set `credit_train` and `credit_test` datasets to be used for training and testing
* Convert the response variable (Label) to a class variable ( instead of 1 and 2 integers), use as.factor
*Use the randomForest::randomForest() function to train a Random Forest model on the credit_train dataset.
The formula used to define the model is the same as for `rpart` -- we want to predict "default" as a function of all the other columns in the training set.
Inspect the model output.



``` 

library(tidyverse)
library(randomForest)

# file to be loaded  
f=file("https://archive.ics.uci.edu/ml/machine-learning-databases/statlog/german/german.data")
  
# load the file   
creditData   <- read.table(f,col.names =paste0("X","1":"21") ,sep=" ")
# select a small subset of columns and  Change the names of the columns 

# colnames(creditData) <- c("chk_acct", "duration", "credit_his", "purpose", "amount", "saving_acct", "present_emp", "installment_rate", "sex", "other_debtor", 
#                            "present_resid", "property", "age", "other_install", "housing", "n_credits", 
#                            "job", "n_people", "telephone", "foreign", "response")

#creditData$response <- german_credit$response - 1
#creditData$response <- as.factor(german_credit$response)

creditsub  <- creditData[,c(2,8,11,13,21)]
names(creditsub) <- c("months_loan_duration","percent_of_income", "years_at_residence","age","default")  
creditsub$default <- ifelse(creditsub$default =="1", "No", "Yes")                  
             
 
smp_size <- f---(0.75 * nrow(creditsub))

## set the seed to make your partition reproducible
set.seed(123)

train_ind <- sa---(seq_len(nrow(cre---)), size = smp_size)

# make a training and test set

credit_test <-  creditsub[-train_ind, ]
credit_train <- creditsub[---, ]

## 
## convert the default variable to factor(category) otherwise the algorithm will consider it as a regression problem 
credit_train$default <- credit_train$default %>% as.factor()
credit_test$default <- credit_test$default %>% as.factor()
## 

set.seed(1)  # Pour le reproductibilité
credit_model <- randomForest(--- ~ ---, data = ---)

# Print  model output                             
print(credit_model) 

```

```{r}
smp_size <- floor(0.75 * nrow(creditsub))

## set the seed to make your partition reproducible
set.seed(123)

train_ind <- sample(seq_len(nrow(creditsub)), size = smp_size)

# make a training and test set
credit_train <- creditsub[train_ind, ]
credit_test <-  creditsub[-train_ind, ]

## 

credit_train$default= credit_train$default %>% as.factor()

credit_test$default= credit_test$default %>% as.factor()
## 

set.seed(1)  # Pour le reproductibilité
credit_model <- randomForest(formula = default ~., data = credit_train)

# Print  model output                             
print(credit_model) 
```

## Evaluate the OOB error



```{r}
# Grab OOB error matrix & take a look

err <- credit_model$err.rate
head(err)
```


```{r}
tail(err, n=1)
```

```{r}
# Print  model output                             
print(credit_model)
```
```{r}
plot(credit_model)
```
